Universals of word order result from optimization of grammars for efficient communication

### 2019-03-31 second-draft/3-paper-names.tex
revising

% ,lin2017critical
\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics \cite{zipf1949human,greenberg1963universals,chomsky1993lectures,...}.
% behaghel1909beziehungen,
% \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three, berwick1984grammatical,hawkins2007processing} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005}.
%Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which certain syntactic universals can be explained through the computational optimization of grammars for efficiency. % information-theoretic

@@

Computational studies

word order difficult, requires large data sets

Drawing on data from 51 languages, we show  that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.
%In this work, we first provide a formalization of efficiency in the domain of word order. %, drawing on information-theoretic notions.
We achieve this using a direct information-theoretic formalization of Zipf's ideas, computationally implemented using state-of-the-art machine learning techniques applied to large-scale text data from 51 languages.

@@

TODO at some point need to foreshadow more of the Greenberg correlations

@@

Consider Figure~\ref{fig:comm}.
Imagine the speaker wants to communicate that Mary saw John.
The message (left) consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.
When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
In natural language, this order is largely determined by the \key{grammar} of the language (CITE):
The speaker will order it as (1) in English, and as (2) in Japanese.

(1)
\begin{dependency}[theme=simple]
    \begin{deptext}
    Mary \& saw \& John \\
    \end{deptext}
    \depedge{2}{1}{nsubj}
    \depedge{2}{3}{obj}
\end{dependency}
(2)
\begin{dependency}[theme=simple]
    \begin{deptext}
    Mary \& John  \& saw \\
    \end{deptext}
    \depedge{3}{1}{nsubj}
    \depedge{3}{2}{obj}
\end{dependency}

The listener receives a sentence -- a linear string of words -- and decodes the underlying message in accordance with the grammar of the language (right).
%Consistent with Shannon's model, we assume a Bayesian listener who decodes dependency structures probabilistically~(Figure~\ref{fig:comm}).
%If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).

@@

The grammars of natural languages provide consistent ordering rules for all syntactic relations.
%All natural languages have some degree of word order regularities that are specified in their grammar.
%For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
%That is, speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
%Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
%Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
%While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
This is illustrated in Figure~\ref{fig:grammars}:
We show how four different grammars order objects, adpositional phrases, and adpositions.
For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.

@@

In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
This idea formalizes Zipf's forces of Unification and Diversification, and overall efficiency can be formalized as a weighted combination of terms representing the amount of information conveyed and the cost of communication:
%The \key{efficiency} of a language is a weighted combination
\begin{equation}\label{eq:efficiency}
	R_{\textit{Eff}} := R_{\textit{Informativity}} - \lambda R_\textit{Cost}
\end{equation}
with an interpolation weight $\lambda \in [0,1)$. %In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification and for robustness to other choices).
Such notions of efficiency have been proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems. %, taking dependency structures as the underlying meanings to be conveyed.
Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship). 
%Word meaning \cite{regier2015word}

@@

@@

>>>>

% ,lin2017critical
\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics \cite{zipf1949human,greenberg1963universals,chomsky1993lectures,...}.
% behaghel1909beziehungen,
% \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005}.
%Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which certain syntactic universals can be explained through the computational optimization of grammars for efficiency. % information-theoretic

@@

In this work, we apply this approach to the problem of explaining Greenberg's word order correlation universals.
Drawing on large sets of text data from 51 languages, we show  that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.
We achieve this using a direct information-theoretic formalization of Zipf's ideas, computationally implemented using state-of-the-art machine learning techniques applied to large-scale text data from 51 languages.

%In this work, we first provide a formalization of efficiency in the domain of word order. %, drawing on information-theoretic notions.

@@

%TODO at some point need to foreshadow more of the Greenberg correlations

@@

Consider Figure~\ref{fig:comm}.
Imagine the speaker wants to communicate that Mary saw John.
The message (left) consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.
When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
In natural language, this order is largely determined by the \key{grammar} of the language (CITE).
The speaker will order it as \emph{Mary saw John} (Figure~\ref{fig:comm}, top) in English, and as \emph{Mary John saw} (Figure~\ref{fig:comm}, bottom) in Japanese.
%
%(1)
%\begin{dependency}[theme=simple]
%    \begin{deptext}
%    Mary \& saw \& John \\
%    \end{deptext}
%    \depedge{2}{1}{nsubj}
%    \depedge{2}{3}{obj}
%\end{dependency}
%(2)
%\begin{dependency}[theme=simple]
%    \begin{deptext}
%    Mary \& John  \& saw \\
%    \end{deptext}
%    \depedge{3}{1}{nsubj}
%    \depedge{3}{2}{obj}
%\end{dependency}
%
The listener receives a sentence -- a linear string of words -- and decodes the underlying message in accordance with the grammar of the language (Figure~\ref{fig:comm}, right).
%Consistent with Shannon's model, we assume a Bayesian listener who decodes dependency structures probabilistically~(Figure~\ref{fig:comm}).
%If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).

@@

%All natural languages have some degree of word order regularities that are specified in their grammar.
%For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
%That is, speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
%Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
%Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
%While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
The grammars of natural languages provide consistent ordering rules for all syntactic relations.
This is illustrated in Figure~\ref{fig:grammars}:
We show how four different grammars order objects, adpositional phrases, and adpositions.
For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.

@@

%In information-theoretic formalizations of Zipf's ideas, the efficiency of a 
In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
This idea formalizes Zipf's forces of Unification and Diversification, and overall efficiency can be formalized as a weighted combination of terms representing the amount of information conveyed and the cost of communication:
%The \key{efficiency} of a language is a weighted combination
\begin{equation}\label{eq:efficiency}
	R_{\textit{Eff}} := R_{\textit{Informativity}} - \lambda R_\textit{Cost}
\end{equation}
with an interpolation weight $\lambda \in [0,1)$. %In all experiments in this paper we use $\lambda=.9$ (see SI appendix section 5 for mathematical justification and for robustness to other choices).
Such notions of efficiency have been proposed in \cite{ferrericancho2002zipf,ferrericancho2007global,futrell2017memory} as a general objective for communicative systems. %, taking dependency structures as the underlying meanings to be conveyed.
Eq.~\ref{eq:efficiency} can also be seen as a simplified form of the Information Bottleneck \cite{tishby1999information}, a general objective function for lossy compression which has recently been applied to explain linguistic phenomena such as color naming systems \cite{zaslavsky2018efficient} (see SI section 4 for the precise relationship). 
%Word meaning \cite{regier2015word}

@@

%CITE other syntactic work? LFG
%TODO the following sentences should go somewhere
%By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
%The dependency structure (including the labels on the arcs) specifies how the meanings of words are combined.
%Therefore, a listener needs to recover the information provided in the dependency structure in order to understand a sentence correctly.

@@

TODO the following paragraphs depend on what visualization we choose
